Sharing and organizing research products as R packages

### 2020-05-22 manuscript/manuscript.Rmd
More revisions to MS

An important component of computational reproducibility is specifying the (software) environment in which the computations were done, or should be done. R packages are a robust solution to this problem, because the package's metadata in the DESCRIPTION file can and should include the R packages and their versions that are required for the project. For example, if the project required using a specific version of R, and a specific version of an R package, this information could be written in the DESCRIPTION file.

@@

There are also at least two other R packages designed to help researchers organize their workflows and facilitate reproducibility and data/code sharing. We highlight these packages here, while also discussing their similarities and differences to our approach. *workflowr* is an R packages that makes it easy to create version controlled and well-organized repositories of research materials, and share them as websites on GitHub [@blischakCreatingSharingReproducible2019; https://github.com/jdblischak/workflowr]. The *workflowr* R package itself is quite mature, and provides helpful tools for compiling many R Markdown documents sequentially, but has a slightly different approach to our proposal, in that workflowr does not itself provide much of a standardized organization for the materials (although it does provide a template). Further, the materials are not organized as an R package, as they would be in the (extended) *vertical* approach. Additionally, *workflowr* does not easily incorporate other types of materials (manuscripts, slides, posters, experiments) in the resulting website, unlike *vertical*. However, it may be slightly easier to use otherwise, for the many useful functions that it provides.

Another approach is contained in the *rrtools* R package [@marwickPackagingDataAnalytical2018; https://github.com/benmarwick/rrtools], which, unlike *workflowr* but like our approach, centers on the R package architecture. Additionally, *rrtools* helps set up a Docker container and continuous integration services for the project, both tools that we hope become more widely adopted in the future. However, it does not help users set up other important research materials as part of the project's materials, such as slides and experiments, nor does it help users share their projects as websites.

>>>>

It is not uncommon to discover breaking changes in previously working code as a result of updates to a computational environment (e.g., a new language version or library dependency). As a result, an important component of computational reproducibility is specifying the (software) environment in which the computations were done, or should be done. In general, there is a continuum of static to dynamic approaches to this problem. A static approach could be a list of the requirements for the computational environment; whereas, a more dynamic approach could be a website or other container that actively runs the required environment (e.g., cloud computing tools such as Binder (https://mybinder.org/), CodeOcean (https://codeocean.com/), or RStudio Cloud (https://rstudio.cloud/)).

For simplicity, we adopt R packages as a robust static solution. For example, an R package's metadata in the DESCRIPTION file can and should include the R packages and their versions that are required for the project. This provides a clear and standardized location for declaring the computational environment  necessary to run the code in the project. Conveniently, end users can opt to install the package with dependencies, which automatically installs libraries listed in the DESCRIPTION file. However, even though the DESCRIPTION may describe the necessary computational environment, an end user may have to take additional steps to create it themselves (e.g., installing R on their computers). Nevertheless, this static approach is not incompatible with dynamic approaches, and advanced users could conceivably adapt vertical websites to link to cloud computing environments that preserve and execute project scripts.

@@

There are also at least two other R packages designed to help researchers organize their workflows and facilitate reproducibility and data/code sharing. We highlight these packages here, while also discussing their similarities and differences to our approach. *workflowr* is an R package that makes it easy to create version controlled and well-organized repositories of research materials, and share them as websites on GitHub (@blischakCreatingSharingReproducible2019; https://github.com/jdblischak/workflowr). The *workflowr* R package itself is quite mature, and provides helpful tools for compiling many R Markdown documents sequentially, but has a slightly different approach to our proposal, in that workflowr does not itself provide much of a standardized organization for the materials (although it does provide a template). Further, the materials are not organized as an R package, as they would be in the (extended) *vertical* approach. Additionally, *workflowr* does not easily incorporate other types of materials (manuscripts, slides, posters, experiments) in the resulting website, unlike *vertical*. However, it may be slightly easier to use otherwise, for the many useful functions that it provides.

Another approach is contained in the *rrtools* R package (@marwickPackagingDataAnalytical2018; https://github.com/benmarwick/rrtools), which, unlike *workflowr* but like our approach, centers on the R package architecture. Additionally, *rrtools* helps set up a Docker container and continuous integration services for the project, both tools that we hope become more widely adopted in the future. However, it does not help users set up other important research materials as part of the project's materials, such as slides and experiments, nor does it help users share their projects as websites.
